Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-22 draft_p2.tex
Fix citations

This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}.

@@

A different approach is the Instanton method~\cite{grafke_instanton_2015}, also relying on the knowledge of an action for the dynamics.

@@

The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.

@@

Whilst Burgers turbulence provids an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.

@@

In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland_computing_2016} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.

@@

Recently, the \cite{gktl} allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}. Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.

@@

A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid. 
At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{ref1,ref2} where the viscosity is artificially enhanced. 
% Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
\\

@@

The flow dynamics are computed by the lattice Boltzmann (LB) method ~\cite{ref1,ref2}. 
While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach. Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic. Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level. This is, in short, the rationale behind the LB approach. In our situation, the LB method has been chosen principally for its outstanding computational efficiency.  
%
More precisely, the fluid is viewed as populations of particles
that collide, redistribute and propagate along the different links of a discrete lattice. 
In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
\[
\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
\]
for the mass density and the fluid momentum respectively. The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.

@@

How rare is a fluctuation $f_d \geq a$ can be determined by computing its \textit{return time} $r(a)$.
It is defined as the average time between two consecutive occurrences of a fluctuation $f_d \geq a$.
Extreme drag fluctuations are \textit{rare events}, that is fluctuations for which $r(a) \gg \tau_c$.
In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang2018}
\begin{equation}
  \label{eq:return_time}
  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
\end{equation}
where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
Figure~\ref{fig:return_time_instant} illustrates the fluctuation amplitude $a$ as a function of the return time $r(a)$, for the drag process $f_d(t)$.
Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang2018}.

@@

The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
This viscous shear is induced by a large vortex detached from the obstacle.
For a more detailed description of such dynamics, see the chapter 3 of ~\cite{ref:these_lestang}.

@@

This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{papier_JSTAT}.
An estimate of this \ac{pdf} is displayed in figure~\ref{fig:pdf_drag}.
Let $X$ be a random variable whose \ac{pdf} is denoted $\mathbb{P}$ and standard deviation $\sigma_X$.
Considering an extreme positive value $a$ of $S_N=\sum_{n=1}{N}X_n$, the probability $p_1$(resp. $p_2$) of case (1) (resp. case (2)) writes:   
\begin{equation}
  \label{eq:indep}
  p_{1}(\sum_{1}^{N} X_n=a) \approx \mathcal{P}\left(\frac{a}{N}\right)^{N} \,\,\, \text{and} \,\,\, p_{2}(\sum_{1}^{N} X_n=a)\approx \mathcal{P}(a)  
\end{equation}
If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
\begin{equation}
  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\sim} C\left(e^{-\alpha \frac{a}{N}}\right)^{N} e^{-\alpha a} = 1.
  \label{eq:ratio_exp}
\end{equation}

@@

% Introduction to AMS
The \acl{ams}~\cite{amspaper} inherits from the basic ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman,rolland_simonnet}.
Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.

% Quick description of TAMS
In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
The \ac{tams} relies on the definition of a time-dependant \textit{score function} $\xi (\mathbf{x}(t),t)$, as well as an iterative procedure starting from an ensemble $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$ of independant trajectories.
Each iteration consists in discarding the trajectories with the lowest maxima of the score function $\xi$ over the whole trajectory of duration $[0;T_a]$.
Discarded trajectories are resampled based on member of the ensemble of trajectories that achieved a higher value of the score function at some point of their history.
As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS}.
For further details about the algorithm and corresponding mathematical results, see~\cite{lestang2018} and references therein.

@@

For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous_varizani,grassberger}.
Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{delMoral_Feynman_Kac_book}.
The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{Giardina2006}.
Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
% It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
% averaging time~\cite{ref_large_dev}.
% By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
% Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
% The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
% The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.

@@

Let us denote formally $\mathbb{P}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right)$ the probability to observe a trajectory $\left\{ x(t)\right\} _{0\leq t\leq T_{a}}$ in the model, and $\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}} \right)$ the probability to observe the same trajectory with the algorithm.
By construction of the algorithm through the weights~\eqref{eq:Weight}, we have
\begin{align}
\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(x(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right).\label{eq:Biased_Path_Approximation}
\end{align}
where the normalisation factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(X(t))dt}\right]$, denoting by $\mathbb{E}_{0}$ the expectation value with respect to $\mathbb{P}_{0}$, and $\underset{N\rightarrow\infty}{\sim}$ means that this is true only asymptotically for large $N$.
The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
Equation~\eqref{eq:Biased_Path_Approximation} is obtained by assuming the mean field approximation
\begin{equation}
R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{0}^{t_{_{1}}}A(X_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{0}^{t_{1}}A(X(t))dt}\right],\label{eq:Mean_Field_Approximation}
\end{equation}
which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{Giardina2006,Giardina2011}:
\begin{equation}
\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(X(t))dt}\right].\label{eq:Estimate_Lambda}
\end{equation}
The validity of the mean field approximation and the fact that the typical relative error due to this approximation is of order $1/\sqrt{N}$ has been proven~\cite{DelMoralBook,DelMoral2013} to be true for a family of rare event algorithms including the one adopted in this paper.

@@

The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$, from the distribution $\mathbb{P}_{k}$.
This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{these_lestang,nemoto_bouchet2016}.
As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
An estimator is then given by
\begin{equation}
\mathbb{E}_{0}\left[O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ X_{n}(t)\right\} _{0\leq t\leq T_{a}}\right)\mbox{e}^{-k\int_{0}^{T_{a}}A(X_{n}(t))dt}\mbox{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
\end{equation}
where the $X_{n}$ are the $N$ backward reconstructed trajectories.
Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
As a result, the effective ensemble of trajectories generated by the \ac{gktl} algorithm contains a larger amount of such events, and empirical estimators such as~\eqref{eq:GK_O_estimator} have a dramatically lower statistical error.

@@

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.

@@

For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
This effect is visible in figure~\ref{fig:IS_GKTL}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.
% how to assess the value of k_max ?
For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

@@

We assessedthe application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.

@@

>>>>

This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice}.

@@

A different approach is the Instanton method~\cite{grafke2015instanton}, also relying on the knowledge of an action for the dynamics.

@@

The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{grafke2015instanton,chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017}.

@@

Although Burgers turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.

@@

In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland2016computing} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.

@@

Recently, the \ac{gktl} algorithm allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}. Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.

@@

A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid. 
At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
% Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
\\

@@

The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}. 
While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach. Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic. Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level. This is, in short, the rationale behind the LB approach. In our situation, the LB method has been chosen principally for its outstanding computational efficiency.  
%
More precisely, the fluid is viewed as populations of particles
that collide, redistribute and propagate along the different links of a discrete lattice. 
In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
\[
\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
\]
for the mass density and the fluid momentum respectively. The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.

@@

How rare is a fluctuation $f_d \geq a$ can be determined by computing its \textit{return time} $r(a)$.
It is defined as the average time between two consecutive occurrences of a fluctuation $f_d \geq a$.
Extreme drag fluctuations are \textit{rare events}, that is fluctuations for which $r(a) \gg \tau_c$.
In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang_computing_2018}
\begin{equation}
  \label{eq:return_time}
  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
\end{equation}
where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
Figure~\ref{fig:return_time_instant} illustrates the fluctuation amplitude $a$ as a function of the return time $r(a)$, for the drag process $f_d(t)$.
Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang_computing_2018}.

@@

The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
This viscous shear is induced by a large vortex detached from the obstacle.
For a more detailed description of such dynamics, see chapter 3 of~\cite{lestang:tel-01974316}.

@@

This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_computing_2018}.
An estimate of this \ac{pdf} is displayed in figure~\ref{fig:pdf_drag}.
Let $X$ be a random variable whose \ac{pdf} is denoted $\mathbb{P}$ and standard deviation $\sigma_X$.
Considering an extreme positive value $a$ of $S_N=\sum_{n=1}{N}X_n$, the probability $p_1$(resp. $p_2$) of case (1) (resp. case (2)) writes:   
\begin{equation}
  \label{eq:indep}
  p_{1}(\sum_{1}^{N} X_n=a) \approx \mathcal{P}\left(\frac{a}{N}\right)^{N} \,\,\, \text{and} \,\,\, p_{2}(\sum_{1}^{N} X_n=a)\approx \mathcal{P}(a)  
\end{equation}
If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
\begin{equation}
  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\sim} C\left(e^{-\alpha \frac{a}{N}}\right)^{N} e^{-\alpha a} = 1.
  \label{eq:ratio_exp}
\end{equation}

@@

% Introduction to AMS
The \acl{ams}~\cite{cerou_adaptive_2007} inherits from the basic ideas of \textit{splitting} algorithms~\cite{KahnHarris1951}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman_multilevel_1999,rolland_statistical_2015}.
Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.

% Quick description of TAMS
In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
The \ac{tams} relies on the definition of a time-dependant \textit{score function} $\xi (\mathbf{x}(t),t)$, as well as an iterative procedure starting from an ensemble $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$ of independant trajectories.
Each iteration consists in discarding the trajectories with the lowest maxima of the score function $\xi$ over the whole trajectory of duration $[0;T_a]$.
Discarded trajectories are resampled based on member of the ensemble of trajectories that achieved a higher value of the score function at some point of their history.
As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS}.
For further details about the algorithm and corresponding mathematical results, see~\cite{lestang_computing_2018} and references therein.

@@

For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous1994go,grassberger2002go}.
Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{giardina_direct_2006}.
Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems~\cite{giardina_direct_2006,tailleur_probing_2007}.
% It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
% averaging time~\cite{ref_large_dev}.
% By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
% Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
% The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
% The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.

@@

Let us denote formally $\mathbb{P}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right)$ the probability to observe a trajectory $\left\{ x(t)\right\} _{0\leq t\leq T_{a}}$ in the model, and $\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}} \right)$ the probability to observe the same trajectory with the algorithm.
By construction of the algorithm through the weights~\eqref{eq:Weight}, we have
\begin{align}
\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(x(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right).\label{eq:Biased_Path_Approximation}
\end{align}
where the normalisation factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(X(t))dt}\right]$, denoting by $\mathbb{E}_{0}$ the expectation value with respect to $\mathbb{P}_{0}$, and $\underset{N\rightarrow\infty}{\sim}$ means that this is true only asymptotically for large $N$.
The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
Equation~\eqref{eq:Biased_Path_Approximation} is obtained by assuming the mean field approximation
\begin{equation}
R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{0}^{t_{_{1}}}A(X_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{0}^{t_{1}}A(X(t))dt}\right],\label{eq:Mean_Field_Approximation}
\end{equation}
which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{giardina_direct_2006,giardina_simulating_2011}:
\begin{equation}
\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(X(t))dt}\right].\label{eq:Estimate_Lambda}
\end{equation}
The validity of the mean field approximation and the fact that the typical relative error due to this approximation is of order $1/\sqrt{N}$ has been proven~\cite{DelMoralBook,DelMoral2013} to be true for a family of rare event algorithms including the one adopted in this paper.

@@

The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$, from the distribution $\mathbb{P}_{k}$.
This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{lestang:tel-01974316,brewer2018efficient}.
As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
An estimator is then given by
\begin{equation}
\mathbb{E}_{0}\left[O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ X_{n}(t)\right\} _{0\leq t\leq T_{a}}\right)\mbox{e}^{-k\int_{0}^{T_{a}}A(X_{n}(t))dt}\mbox{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
\end{equation}
where the $X_{n}$ are the $N$ backward reconstructed trajectories.
Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
As a result, the effective ensemble of trajectories generated by the \ac{gktl} algorithm contains a larger amount of such events, and empirical estimators such as~\eqref{eq:GK_O_estimator} have a dramatically lower statistical error.

@@

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{touchette_large_2009}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{touchette_large_2009}.

@@

For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
This effect is visible in figure~\ref{fig:IS_GKTL}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
See~\cite{nemoto2017finite,Guevara_Hidalgo_2018} for further details.
% how to assess the value of k_max ?
For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

@@

We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.

@@

\bibliographystyle{unsrt}
\bibliography{draft_p2}
